Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-21 mard-survey.tex
2012.05.18 (Hüseyin) Updating table to be more compatible with the text. Figures added/updated.

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella-like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, the search is performed
on a BFS tree with multiple preset depths. The depth limit is iteratively
increased by the source node for each query based on the quality of the results.
The source node, after examining the results, may issue a new request by
increasing the depth limit, which will trigger the nodes at the last depth
level to resume the search. The iterative approach avoids to start the search
from scratch at each iteration and reduces the load of the nodes on the upper
levels of the tree. The major drawback of \textit{iterative deepening} is the
delay between successive iterations, as the source node needs to examine the
results at each iteration before deciding to quit or resume the query.  The
\textit{directed BFS} tries to avoid this delay by forwarding the query messages
only to a selected set of neighbours, in which the selection criteria varies
from the number of results received previously, distance in terms of hops,
bandwidth, or the query load of the neighbour. In \textit{local indices}, each
node sustains a local data index of the nodes within a radius of $r$ hops from
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic.  \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

@@

Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for creating self-adapting overlay networks, that can achieve application-layer multicast communication without requiring IP multicast infrastructure at the network layer. IP multicast is the term reffering to the method of sending IP datagrams to a group of receivers in a single transmission. Even though the method is available some years now, IP multicast has not taken off as anticipated since many believe that it violates the stateless design of the current Internet. For this reason it is widely deployed in contained environments like enterprises (i.e. IPTV applications - distance learning, video conferencing), commercial stock exchanges, and multimedia content delivery networks. Narada tries to break these constraints. Although it was not originally designed as a P2P file sharing protocol, it has become one of the pioneering works in the overlay networks area. The authors claim that, although Narada is not as efficient as IP multicast, still it provides reasonable performance, without the additional cost of implementing multicast services on the current Internet infrastructure. Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}. They also keep the
graph and the trees dynamically updated while nodes continue to join and leave the network. Moreover, the protocol tries to ease the physical link stress, the overall resource usage and the relative delay among end systems. Unfortunately, the most imporpant limitation of Narada is that although it works reasonably well for small groups, it does not scale well for larger networks, therefore it is not suitable for, potentially very large, P2P file sharing application networks.

@@

Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes across the Internet. In order to detect locality, nodes use network latency (round trip
time) as a measurement technique. The network latency, even though not always
accurate, is selected in this work due to its non-intrusive, transparent and
easy to apply behaviour. In order for the binning to work, few well known anchor servers with known physical locations on the Internet. Authors estimate that a number of 12 such servers can prove sufficient for this. The nodes measure their distance to all these servers and record the latency values for each. Later, the tuple of these latency values represents the bin of each node. Nodes with similar latency values are considered close to each other, or at least within the same autonomous system. The operation of the method is independent of the model incorporated by the overlay network and thus it can be applied with no significant changes to both structured and unstructured P2P systems. \cite{ratnasamy_binning_2002} has the detailed description of such algorithm for both architectures. Landmark binning is also considered as a good candidate for content distribution networks in particular. The major disadvantage of landmark
binning is to install and maintain landmark servers on different autonomous systems, all over the world.  A typical P2P network usually has a couple of million nodes
connected at any given time, thus rendering landmark servers an important parameter when designing for large scales.  The authors claim that latency estimation does not drain
much from the network resources but in case scalability problems arise, they claim the solution is replacing single landmark servers with clusters of servers
within the same physical area. However, this approach does not reduce the
possibility of excessive network traffic flow through these landmark servers. One other
possible problem is the incorrect binning caused by the use of inaccurate methods for delay measurement, since tracking down, for example, network latencies is far from proven to be safe for location estimation.

@@

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{tabular}{
\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
\textbf{Narada} & Decentralized unstructured & \textbf{Overlay optimization
based}. Creates a mesh (richer connected graph) and builds minimum spanning
trees on this mesh & & Yes. Supports incremental evolving & & Small and sparse
groups & 2000  \\

\hline
\textbf{Gia} & Decentralized unstructured & \textbf{Forwarding based} Replaces
Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
dynamic topology adaptation protocol &
 Gnutella & Yes & & Better than Gnutella & 974  \\

\hline
\textbf{Adaptive Overlay Topology Optimization} & Decentralized unstructured & \textbf{Overlay optimization
based}. Creates overlay multicast tree with Selective Flooding protocol&
Gnutella & Yes, using Active Topology protocol & & Better than Gnutella & 36 \\

\hline
\textbf{Location-aware Topology Matching} & Decentralized unstructured &
\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
connection cutting}, and \textit{source peer probing}. & Gnutella & Yes & & Better than Gnutella & 204 \\

\hline
\textbf{Replication Strategies in Unstructured P2P Networks} & Decentralized unstructured &
\textbf{Cache Based}. Uses uniform, proportional and square root allocation
strategies to replicate data. & Gnutella & Yes & & Better than Gnutella & 580 \\

\hline
\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} & Decentralized unstructured &
\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella &
Yes & & Better than Gnutella & 199 \\

\hline
\textbf{Improving search in P2P networks} & Decentralized unstructured &
\textbf{Forwarding Based}. Uses \textit{iterative deepening}, \textit{directed
BFS}, and \textit{local indices} to improve efficiency. & Gnutella & Yes & & Better than Gnutella & 380 \\

\hline
\textbf{Distributed Cycle Minimization Protocol} & Decentralized unstructured &
\textbf{Forwarding based} Uses a decentralized cycle elimination protocol  &  & Yes &  &  & 9 \\

\hline
\textbf{Scalable Bipartite Overlay} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses bipartite partition graph and builds
local minimum spanning trees  & Gnutella & Yes &  & Better than Gnutella & 77 \\

\hline
\textbf{Adaptive Connection Establishment} & Decentralized unstructured &
\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
minimum spanning trees and perform local optimizations & Adaptive Overlay
Topology Optimization (AOTO), Gnutella & Yes &  & Better than Gnutella & 110 \\

\hline
\textbf{Hops Adaptive Neighbour Discovery} &  &
 &  &  &  &  & 1 \\

\hline
\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
distances and selects neighbours  & Gnutella & Yes &  &  & 36 \\

\hline
\textbf{mOverlay} & Decentralized unstructured &
\textbf{Overlay optimization based} Uses dynamic landmarks to find node locality
& & Yes &  & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols & 123 \\

\hline
\textbf{Distributed Domain Name Order (DDNO)} & Decentralized unstructured &
\textbf{Overlay optimization based} Connects half of the nodes connections to
the nodes in the same domain and the other half to random nodes, therefore
supports locality and topological connection  & & Yes &  & Yes, by using super
peers & 5 \\

\hline
\textbf{Peer-exchange Routing Optimization Protocols} & Works with both decentralized
structured and unstructured &
\textbf{Overlay optimization based} Optimizes overlay by the exchange of
neighbors among peers  & & Yes &  & Yes & 13 \\

\hline
\textbf{MAY OMIT- T2MC} & Decentralized unstructured &
\textbf{Cluster based---{\sethlcolor{yellow}\hl{ IS IT ALSO PROXIMITY BASED?}}} Uses traceroute results for clustering the
nodes  & &  &  &  & 1 \\

\hline
\textbf{Unnamed-unstructured} & Decentralized unstructured &
\textbf{Overlay optimization based} Minimizes the communication delay and
maximizes the broadcasting range & & Yes &  & Better than THANCS and mOverlay & 7\\

\hline
\textbf{Landmark Binning} & Can work with both decentralized structured and unstructured &
\textbf{Proximity based, landmark binning} Uses network latency to partition
nodes into bins & & Yes &  &  & 817 \\

\hline
%\end{tabular}
\end{longtable}
\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\end{figure}

@@

\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. This global view of the state is made available to all
nodes in order to help them find the best way to route their messages.
\textit{Global Soft-State} operates in two main stages, generation and use of
the proximity information. For generating the proximity information a hybrid
approach is proposed, which uses landmark clustering as a preprocessing step in
order to select a number of potential nearest neighbour candidates and then
refine the selection by incorporating a round-trip-time-based scheme to ultimately choose the
closest node. For using the proximity information, the algorithm chooses a
different path from the classic gossiping approaches for constructing and
maintaining the overlay. It is based on landmark clustering for strategic
placement of proximity information on the overlay enabling any node to access
such information using a landmark number that reflects its physical position in
the network. For various logical regions\footnote{This might be a high-order
zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a
particular prefix in overlays such as Pastry.} maps of physical information are
built and published where each node may appear in a maximum of $log\left( N
\right)$ such maps. To dynamically adapt to changing network conditions, a node
subscribes to relevant \emph{soft states} that utilize a notification system in
order to initialize any necessary neighbour re-selection.

@@

\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
tries to solve the topology mismatch problem by mapping the overlay network as
close as possible to the physical one. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, \textit{SAT-Match} peers can do
selective jumps to adjust their positioning in the DHT, if it reduces the stretch
of its one-hop neighbourhood. The stretch is defined as the ratio between the
average logical and physical latency of a link. It is reported that, due to the
selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in link
stretch, and when used with the \textit{Landmark Binning} approach (see Sec.
\ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
For dynamic environments, with frequent node arrival and leave,
\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
adaptation mechanism and selective jumps.

@@

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Arch.} & \textbf{Overlay structure} & \textbf{Base protocol} &
\textbf{Dynamic update} & \textbf{Runtime} & \textbf{Scalability} & \textbf{cites}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot

\hline
\textbf{Global Softstate} & decentralized structured &
\textbf{Proximity based, landmark clustering} Uses first landmark clustering
then measures RTTs to identify close nodes & & Yes &  &  & 195 \\

\hline
\textbf{Mithos} & decentralized structured &
\textbf{Proximity based} Uses nodes as topology landmarks and directed
incremental probing to optimize topology & & Yes &  & Scales well as all
operations are local ??? & 172 \\

\hline
\textbf{Self-Adaptive Topology Matching} & decentralized structured &
\textbf{Proximity based} Uses lightweight probing and
selective jumps to optimize the topology & CAN & Yes &  & Better than Mithos
& 40 \\

\hline
\textbf{Delay Aware P2P System} & &
\textbf{} & &  &  &  & 1 \\

\hline
\textbf{VERSION OF CHORD - DHT-PNS} & decentralized structured &
\textbf{Proximity based} Uses Proximity Neighbour Selection and the Vivaldi
system & Chord  & Yes &  &  & 5 \\

\hline
\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} & decentralized structured &
\textbf{Proximity based}  & Chord  & Yes &  &  & 0 \\

\hline
\textbf{LAPTOP} & decentralized structured &
\textbf{Geographic layout based} Hierarchical overlay structure  &   & Yes &  & routing path length $\log{_d N}$, 
join/leave overhead $d\log{_d N}$& 4 \\

\hline
\textbf{IP-Based Clustering} & decentralized structured &
\textbf{Proximity based} Proximity neighbour selection based on longest common
prefix of IP addresses &   & Yes &  &  & 1 \\

\hline
\textbf{CHOord considering Proximity on IPv6} & decentralized structured &
\textbf{Proximity based} Uses IPv6 address format to provide proximity &  Chord
& Yes &  & Better than Chord & 1 \\

\hline
\textbf{Proximity in Kademlia} & decentralized structured &
\textbf{Proximity based} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
to Kademlia & Kademlia & Yes &  &  & 13 \\

\hline
\textbf{Cone} & decentralized structured &
\textbf{Proximity based} Uses proximity neighbour selection (PNS) & Chord & Yes
&  & Better than Chord & 3 \\

\hline
\textbf{Dynamo} &  & &  & 
&  & & 3 \\

\hline
\textbf{MAY OMIT-BADLY WRITTEN-PChord} & decentralized structured &
\textbf{Proximity based}  &  Chord
& Yes &  & & 16 \\

\hline
\textbf{AChord} & &  &  
&  &  & & 7 \\

\hline
\textbf{Chord6} & decentralized structured &
\textbf{Proximity based} Uses IPv6 hierarchical address format to cluster
topologically close nodes & Chord & Yes
&  &  & 14 \\

\hline
\end{longtable}
\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\label{fig:struct_compare_table}
%\end{figure}

>>>>

\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/iterative_deepening}
\caption{Iterative deepening with three levels.}
\label{fig:iterative_deepening}
\end{figure*}

\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/directed_bfs}
  \includegraphics[scale=0.4]{img/algorithms/local_indices}
\caption{(left) directed BFS. (right) local indices with radius size equal to $2$.}
\label{fig:directed_bfs}
\end{figure*}

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella-like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, as shown in Figure \ref{fig:iterative_deepening},
the search is performed
on a BFS tree with multiple preset depths. The depth limit is iteratively
increased by the source node for each query based on the quality of the results.
The source node, after examining the results, may issue a new request by
increasing the depth limit, which will trigger the nodes at the last depth
level to resume the search. The iterative approach avoids to start the search
from scratch at each iteration and reduces the load of the nodes on the upper
levels of the tree. The major drawback of \textit{iterative deepening} is the
delay between successive iterations, as the source node needs to examine the
results at each iteration before deciding to quit or resume the query.  The
\textit{directed BFS}, as shown in Figure \ref{fig:directed_bfs}(left), tries to avoid this delay by forwarding the query messages
only to a selected set of neighbours, in which the selection criteria varies
from the number of results received previously, distance in terms of hops,
bandwidth, or the query load of the neighbour. In \textit{local indices}, as shown in Figure \ref{fig:directed_bfs}(right), each
node sustains a local data index of the nodes within a radius of $r$ hops from
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic.  \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

@@

\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/narada}
  \includegraphics[scale=0.4]{img/algorithms/narada2}
  \includegraphics[scale=0.4]{img/algorithms/narada3}
\caption{(left) underlying network with edge costs. (middle) Peer A sends a message to the rest using regular broadcast. (right) Peer A sends a message to the rest using end system multicast.}
\label{fig:narada}
\end{figure*}

Narada \cite{chu_esm_2000,chu_esm_2001,chu_esm_2002} is a generic protocol for creating self-adapting overlay networks, that can achieve application-layer multicast communication without requiring IP multicast infrastructure at the network layer. IP multicast is the term reffering to the method of sending IP datagrams to a group of receivers in a single transmission. Even though the method is available some years now, IP multicast has not taken off as anticipated since many believe that it violates the stateless design of the current Internet. For this reason it is widely deployed in contained environments like enterprises (i.e. IPTV applications - distance learning, video conferencing), commercial stock exchanges, and multimedia content delivery networks. Narada tries to break these constraints. Although it was not originally designed as a P2P file sharing protocol, it has become one of the pioneering works in the overlay networks area. The authors claim that, although Narada is not as efficient as IP multicast, still it provides reasonable performance, without the additional cost of implementing multicast services on the current Internet infrastructure. Additionally, authors realised the inefficiency caused by the topology mismatch problem and tried to encounter it by building a richer connected graph, called a mesh, and building per source minimum spanning trees \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction} as shown in Figure \ref{fig:narada}. They also keep the
graph and the trees dynamically updated while nodes continue to join and leave the network. Moreover, the protocol tries to ease the physical link stress, the overall resource usage and the relative delay among end systems. Unfortunately, the most imporpant limitation of Narada is that although it works reasonably well for small groups, it does not scale well for larger networks, therefore it is not suitable for, potentially very large, P2P file sharing application networks.

@@

\begin{figure*}
\centering
  \includegraphics[scale=0.4]{img/algorithms/landmark_binning}
\caption{Example 2D coordinate overlay and a sample routing path from node D to (x,y).}
\label{fig:landmark_binning}
\end{figure*}

Landmark Binning \cite{ratnasamy_binning_2002} is an approach to partition close
by nodes into bins based on their distance to well known anchor nodes across the Internet, as shown in Figure \ref{fig:landmark_binning}. In order to detect locality, nodes use network latency (round trip
time) as a measurement technique. The network latency, even though not always
accurate, is selected in this work due to its non-intrusive, transparent and
easy to apply behaviour. In order for the binning to work, few well known anchor servers with known physical locations on the Internet. Authors estimate that a number of 12 such servers can prove sufficient for this. The nodes measure their distance to all these servers and record the latency values for each. Later, the tuple of these latency values represents the bin of each node. Nodes with similar latency values are considered close to each other, or at least within the same autonomous system. The operation of the method is independent of the model incorporated by the overlay network and thus it can be applied with no significant changes to both structured and unstructured P2P systems. \cite{ratnasamy_binning_2002} has the detailed description of such algorithm for both architectures. Landmark binning is also considered as a good candidate for content distribution networks in particular. The major disadvantage of landmark
binning is to install and maintain landmark servers on different autonomous systems, all over the world.  A typical P2P network usually has a couple of million nodes
connected at any given time, thus rendering landmark servers an important parameter when designing for large scales.  The authors claim that latency estimation does not drain
much from the network resources but in case scalability problems arise, they claim the solution is replacing single landmark servers with clusters of servers
within the same physical area. However, this approach does not reduce the
possibility of excessive network traffic flow through these landmark servers. One other
possible problem is the incorrect binning caused by the use of inaccurate methods for delay measurement, since tracking down, for example, network latencies is far from proven to be safe for location estimation.

@@

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{tabular}{
%\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Unstructured Algorithms} \label{fig:unstruct_compare_table} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
 \textbf{Scalability}\\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} &  \textbf{Overlay structure} & \textbf{Base protocol} &
 \textbf{Scalability}\\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot
\textbf{Narada} & \textbf{Overlay optimization
based}. Creates a mesh (richer connected graph) and builds minimum spanning
trees on this mesh & & Small and sparse groups \\

\hline
\textbf{Gia} & \textbf{Broadcast based} Replaces
Gnutella flooding with random walk, and introduces KaZaA style supernodes. Uses
dynamic topology adaptation protocol &
 Gnutella &  Better than Gnutella  \\

\hline
\textbf{Adaptive Overlay Topology Optimization} & \textbf{Overlay optimization
based}. Creates overlay multicast tree with Selective Flooding protocol&
Gnutella &  Better than Gnutella \\

\hline
\textbf{Location-aware Topology Matching} & 
\textbf{Overlay Optimization Based}. Uses \textit{TTL2-detector flooding}, \textit{low productive
connection cutting}, and \textit{source peer probing}. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Replication Strategies in Unstructured P2P Networks} &
\textbf{Cache Based}. Uses uniform, proportional and square root allocation
strategies to replicate data. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Tracing a large-scale Peer to Peer System: an hour in the life of Gnutella.} &
\textbf{Cache Based}. Proposes a caching algorithm based on the traces of the Gnutella traffic & Gnutella & Better than Gnutella \\

\hline
\textbf{Improving search in P2P networks} &
\textbf{Broadcast Based}. Uses \textit{iterative deepening}, \textit{directed
BFS}, and \textit{local indices} to improve efficiency. & Gnutella &  Better than Gnutella \\

\hline
\textbf{Distributed Cycle Minimization Protocol} & 
\textbf{Broadcast based} Uses a decentralized cycle elimination protocol  &  &  \\

\hline
\textbf{Scalable Bipartite Overlay} & 
\textbf{Overlay optimization based} Uses bipartite partition graph and builds
local minimum spanning trees  & Gnutella  & Better than Gnutella \\

\hline
\textbf{Adaptive Connection Establishment} & 
\textbf{Overlay optimization based} Forms Neighbour Cost Tables, builds local
minimum spanning trees and perform local optimizations & Adaptive Overlay
Topology Optimization (AOTO), Gnutella & Better than Gnutella \\

\hline
\textbf{Hops Adaptive Neighbour Discovery} &  & &  \\

\hline
\textbf{Two-Hop-Away Neighbour Comparison and Selection (THANCS)} &
\textbf{Overlay optimization based} Uses piggybacking to discover neighbour
distances and selects neighbours  & Gnutella  & \\

\hline
\textbf{mOverlay} &\textbf{Landmark based proximity} Uses dynamic landmarks to find node locality
& & Due to dynamic landmarks and grouping, more scalable than tree-based or mesh-based protocols \\

\hline
\textbf{Distributed Domain Name Order (DDNO)} &
\textbf{Overlay optimization based} Connects half of the nodes connections to
the nodes in the same domain and the other half to random nodes, therefore
supports locality and topological connection  &  & Yes, by using super
peers \\

\hline
\textbf{Peer-exchange Routing Optimization Protocols} & \textbf{Overlay optimization based} Optimizes overlay by the exchange of
neighbors among peers  & Can work with both decentralized structured and
unstructured architecture & Yes \\

\hline
\textbf{MAY OMIT- T2MC} &
\textbf{Overlay optimization based} Uses traceroute results for clustering the
nodes  & & \\

\hline
\textbf{Unnamed-unstructured} & 
\textbf{Overlay optimization based} Minimizes the communication delay and
maximizes the broadcasting range & & Better than THANCS and mOverlay \\

\hline
\textbf{Landmark Binning} & \textbf{Landmark based proximity} Uses network latency to partition
nodes into bins & Can work with both decentralized structured and unstructured architecture & \\

\hline
%\end{tabular}
\end{longtable}
%\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\end{figure}

@@

\begin{figure*}
\centering
  \includegraphics[scale=0.8]{img/algorithms/global_softstate}
\caption{(left) Position of $p$ in the space of $3$ landmarks. (right) Position of $p$ on the map.}
\label{fig:global_softstate}
\end{figure*}

\textit{Global Soft-State} \cite{xu_globstate_2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. Figure \ref{fig:global_softstate} presents the landmark space and the mapping of the position of the nodes. This global view of the state is made available to all
nodes in order to help them find the best way to route their messages.
\textit{Global Soft-State} operates in two main stages, generation and use of
the proximity information. For generating the proximity information a hybrid
approach is proposed, which uses landmark clustering as a preprocessing step in
order to select a number of potential nearest neighbour candidates and then
refine the selection by incorporating a round-trip-time-based scheme to ultimately choose the
closest node. For using the proximity information, the algorithm chooses a
different path from the classic gossiping approaches for constructing and
maintaining the overlay. It is based on landmark clustering for strategic
placement of proximity information on the overlay enabling any node to access
such information using a landmark number that reflects its physical position in
the network. For various logical regions\footnote{This might be a high-order
zone in the eCAN\cite{xu_ecan_2002} context or a set of nodes sharing a
particular prefix in overlays such as Pastry.} maps of physical information are
built and published where each node may appear in a maximum of $log\left( N
\right)$ such maps. To dynamically adapt to changing network conditions, a node
subscribes to relevant \emph{soft states} that utilize a notification system in
order to initialize any necessary neighbour re-selection.

@@

\begin{figure*}
\centering
  \includegraphics[scale=0.6]{img/algorithms/SAT_match}
  \hspace{3ex}
  \includegraphics[scale=0.6]{img/algorithms/SAT_match2}
\caption{(left) Example physical topology of nodes. (middle) A jumps to B in order to match the logical topology to the physical. (right) State of the logical topology after the jump.}
\label{fig:SAT_match}
\end{figure*}

\textit{SAT-Match} \cite{ren_satmatch_2004} is a protocol that specifically
tries to solve the topology mismatch problem by mapping the overlay network as
close as possible to the physical one. Even though no theoretical proof is
given, valid results are obtained and demonstrated through experimentations.
Similar to the other approaches in the proximity neighbour selection,
\textit{SAT-Match} uses probing to detect close by peers and obtain proximity
information. However, as an additional feature, \textit{SAT-Match} peers can do
selective jumps to adjust their positioning in the DHT, if it reduces the stretch
of its one-hop neighbourhood. Figure \ref{fig:SAT_match} demonstrates the selective jump on an example topology.
The stretch is defined as the ratio between the
average logical and physical latency of a link. It is reported that, due to the
selective jumps, the \textit{SAT-Match} achieves $40\%$ reduction in link
stretch, and when used with the \textit{Landmark Binning} approach (see Sec.
\ref{sec:landmark_binning}), the reduction rate increases up to $60\%$.
For dynamic environments, with frequent node arrival and leave,
\textit{SAT-Match} scales much better than \textit{Mithos}, due to its self
adaptation mechanism and selective jumps.

@@

%\begin{figure}[h!]
\hspace{-3ex}
\begin{center}
\footnotesize
%\begin{landscape}
\begin{longtable}{
|>{\columncolor[gray]{.7}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.4\columnwidth}
|>{\columncolor[gray]{.8}}m{0.2\columnwidth}
|>{\columncolor[gray]{.9}}m{0.2\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.9}}m{0.1\columnwidth}
%|>{\columncolor[gray]{.8}}m{0.1\columnwidth}
|}
\caption{Decentralized Structured Algorithms}\label{fig:struct_compare_table}\\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
\hline
\endfirsthead
\multicolumn{4}{c}%
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline
\rowcolor[gray]{.5}
\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Base protocol} & \textbf{Scalability} \\
\hline
\endhead
\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
\endfoot
\hline
\endlastfoot

\hline
\textbf{Global Softstate} & 
\textbf{Geographic layout} Uses first landmark clustering
then measures RTTs to identify close nodes &  &  \\

\hline
\textbf{Mithos} &
\textbf{Geographic layout} Uses nodes as topology landmarks and directed
incremental probing to optimize topology & & Scales well as all
operations are local ??? \\

\hline
\textbf{Self-Adaptive Topology Matching} &
\textbf{Proximity neighbour selection} Uses lightweight probing and
selective jumps to optimize the topology & CAN &  Better than Mithos \\

\hline
\textbf{Delay Aware P2P System} & \textbf{} & & \\

\hline
\textbf{VERSION OF CHORD - DHT-PNS} &
\textbf{Proximity neighbour selection} Uses Proximity Neighbour Selection and the Vivaldi
system & Chord  &  \\

\hline
\textbf{MAY OMIT- VERSION OF CHORD - Quasi-Chord} &
\textbf{Proximity based}  & Chord  &  \\

\hline
\textbf{LAPTOP} & 
\textbf{Geographic layout} Hierarchical overlay structure  &  & routing path length $\log{_d N}$, 
join/leave overhead $d\log{_d N}$ \\

\hline
\textbf{IP-Based Clustering} & 
\textbf{Proximity neighbour selection} Proximity neighbour selection based on longest common
prefix of IP addresses &    &  \\

\hline
\textbf{CHOord considering Proximity on IPv6} & 
\textbf{Proximity routing} Uses IPv6 address format to provide proximity &  Chord
 & Better than Chord \\

\hline
\textbf{Proximity in Kademlia} & 
\textbf{Proximity routing} Applies  proximity neighbour selection (PNS) and proximity route selection (PRS)
to Kademlia & Kademlia &   \\

\hline
\textbf{Cone} & 
\textbf{Proximity neighbour selection} Uses proximity neighbour selection (PNS) & Chord  & Better than Chord \\

\hline
\textbf{Dynamo} &  &  &  \\

\hline
\textbf{MAY OMIT-BADLY WRITTEN-PChord} & 
\textbf{Proximity based}  &  Chord  & \\

\hline
\textbf{AChord} &  & & \\

\hline
\textbf{Chord6} & 
\textbf{Proximity routing} Uses IPv6 hierarchical address format to cluster
topologically close nodes & Chord  &  \\

\hline
\end{longtable}
%\end{landscape}
\end{center}
\vspace{-2.5ex}
\vspace{-2.5ex}
%\label{fig:struct_compare_table}
%\end{figure}
